### 2017-09-13 papers/BijectiveLensSynth/popl_18_rebuttal.txt

Formally, they are incomparable: each can synthesize some transformations
that the other cannot.

Since we restrict to bijections, we are unable to express functions that
don't preserve data, such as
  - mapping from a format with whitespace to one without whitespace,
  - turning a mix of uppercase and lowercase characters into purely
    uppercase characters, and
  - extracting a file name from a file path.
FlashFill can express all these.

Conversely, our "Optician" algorithm can solve some classes of problems that
existing synthesis systems cannot.  In particular, techniques that only use
input-output examples are unable to synthesize programs with nested loops.

One more technical note about point: FlashFill's DSL does permit nested
loops, but for efficiency reasons the synthesis algorithm only synthesizes
programs with no nesting (this is discussed at the end of section 4.4 of
"Automating String Processing in Spreadsheets Using Input-Output
Examples", POPL '11).  While a single nesting of loops seems to be
sufficient FlashFIll's intended uses, our benchmark suite (based on
real-world uses of the Augeas tool) includes many more complex formats and
transformations.  In particular, these formats tend to have deep layers of
nested loops, with complex transformations at each layer.
FlashFill cannot handle such examples.

[BCP: The last sentence or two are not convincing yet!

@@

                                -------------

     - What modifications did we need to make to examples in the Augeas
       benchmark suite to make them bijective?

Many reviewers expressed concern that the changes made to the formats in our
benchmark suite to make them bijective were major and time consuming. The
changes we made to the benchmark suite were of two forms:
1) Whitespace was present in one format but not the other.
2) Useful information was projected away going from one format to the other.

@@

Even the relatively simple lens that permuts the 3 elements needed for
single_author_convert is fairly complex.  The complexity becomes even more
apparent when working with large lenses comprised of many sub-lenses (though
for brevity we give here the more minimal example above).  Lenses provide
great power with their invertibility guarantees, but they come at the cost
of thinking about many fiddly details when writing the terms.  Worrying
about these fiddly details while ALSO thinking about unambiguity
restrictions is a very difficult task!

Optician is aimed at different audiences than existing string transformation
tools.  While existing tools allow the average end-user to generalize their
examples to programs, Optician allows programmers to get the benefits of a
powerful language while avoiding much of the complexity of programming in
that language.  This paradigm allows for more complex programs to be
synthesized, the types of programs that necessitate programming in such a
DSL, but also provide a burden on the user.  However, this burden is already
one incured by programmers using best practices, so we are merely saving
them the time spent on tedious work.

===========================================================================
===========================================================================
Detailed responses to individual questions in the reviews follow...

===========================================================================
Review #18A

@@

We can make sure we can provide at a description of the sizes of specs and
lenses in the final version of the paper.

@@

With minimization, the title field transformation would appear as:

const("<Field Id=2>","")
  . (Id("")
       | (Id(text_char) . iterate(Id(text_char)) . const("", ",")))
  . const("</Field>",""))

In particular, the useless concatenations of const("","") are removed and
the field boilerplate is removed from being on both sides of the disjunct.
Furthermore, when const(s,s) is present, it is converted into Id(s).

    - Eval the text says Full synthesized 49 benchmarks, but the plot in fig 8
      only shows 38/39?

### 2017-09-13 papers/BijectiveLensSynth/popl_18_rebuttal.txt

Formally, they are incomparable: each can synthesize some transformations
that the other cannot.

Since we restrict to bijections, we are unable to express functions that
don't preserve data, such as
  - mapping from a format with whitespace to one without whitespace,
  - turning a mix of uppercase and lowercase characters into purely
    uppercase characters, and
  - extracting a file name from a file path.
FlashFill can express all these.

Conversely, our "Optician" algorithm can solve some classes of problems that
existing synthesis systems cannot.  In particular, techniques that only use
input-output examples are unable to synthesize programs with nested loops.

One more technical note about point: FlashFill's DSL does permit nested
loops, but for efficiency reasons the algorithm only synthesizes
programs with no nesting (this is discussed at the end of section 4.4 of
"Automating String Processing in Spreadsheets Using Input-Output
Examples", POPL '11).  While a single nesting of loops seems to be
sufficient for FlashFill's intended uses, our benchmark suite (based on
real-world uses of the Augeas tool) includes many more complex formats and
transformations.  In particular, these formats tend to have deep layers of
iterated data, with complex transformations at each layer, requiring the
synthesis of many layers of nested loops.  FlashFill cannot handle such
examples.

@@

                                -------------

     - What modifications did we need to make to examples in the Augeas and
       FlashFill benchmark suite to make them bijective?

Many reviewers expressed concern that major changes were added to the formats in
our benchmark suite to make them bijective. We made changes to the benchmark
suite to handle two forms of non-bijectivity:
1) Whitespace was present in one format but not the other.
2) Useful information was projected away going from one format to the other.

@@

Category 1 was by far the most common, applying to many of the Augeas
examples.  To handle this, we added additional whitespace to the other format,
where typically whitespace is unnecessary.

Category 2 was less prevalent, but still present, typically in the FlashFill
examples.  To address this category, we added additional information to the end
of the target format containing the projected information.

     - The restriction to bijective lenses is not fully justified

There are still many formats kept in bijective correspondence with each other,
like those present at https://www.data.gov/ .

There are lens theories developed to handle synchronization tasks for data
formats not kept in bijective correspondence (quotient lenses to handle Category
1 non-bijectivity, and symmetric lenses to handle Category 2 non-bijectivity).
These theories have a bijective core, surrounded by either canonizers (for
quotient lenses) or projections (for symmetric lenses).  We believe that our
system can be used as a drop-in component for synthesis algorithms on these more
complex lens structures.

These justifications can be added to the paper.

@@

      let bibtex_to_endnote_au =
        synth
          BIBTEXAUTHORINFO <=> TAGGEDAUTHORDEFNS

      let bibtex_to_endnote_au_synthesized_result
        : (lens in BIBTEXAUTHORINFO <=> TAGGEDAUTHORDEFNS) =
        del "author={"
          . ins "au - "
          . lens_swap (NAME . del ",") (lens_swap WSP NAME)* 
          . (del " and "
               . ins "\n au - "
               . lens_swap (NAME . del ",") (lens_swap WSP NAME)* )* 
          . del "}"

      let bibtex_to_endnote_au_human_written
        : (lens in BIBTEXAUTHORINFO <=> TAGGEDAUTHORDEFNS) =
        let single_author_convert =
          ins " au - "
            . lens_swap
                (NAME . del ",")
                (lens_swap WSP NAME)*
        in
        del "author={"
          . single_author_convert
          . (del " and "
               . ins "\n"
               . single_author_convert)* 
          . del "}"

Even the relatively simple lens that permutes the 3 elements needed for
single_author_convert is fairly complex.  The complexity becomes even more
apparent when working with large lenses comprised of many sub-lenses (though
for brevity we give here the more minimal example above).  Lenses provide
great power with their invertibility guarantees, but they come at the cost
of thinking about many fiddly details when writing the terms.  Worrying
about these fiddly details while ALSO thinking about unambiguity
restrictions is a very difficult task!

@@

See above.

                                -------------

     - I also have a concern: since the approach is
       limited to bijective lenses, but many of the test examples were not
       bijective, how much work was it to modify the examples
       and how easy is it to manually hack the resulting bijective
       lenses to make them work on the real data?  Is there any chance of
       automating this or generalizing to handle non-bijective lenses
       (perhaps with this algorithm as a subroutine)?

For Category 1 non-bijectivity, instead of changing the formats, we could have
placed canonizers on the edges of formats with large amounts of whitespace.
Canonizers can normalize an arbitrarily long sequence of whitespace to a single
space.  Canonizers also can be used to handle casing issues: normalizing a
sequence of uppercase and lowercase characters into purely uppercase characters.
In this manner, whitespace (or casing) information can be normalized, and no
longer would need to be propagated to the target format. We have recently
developed a framework to annotate a regular expression with canonizing
information, allowing for a single term to represent an arbitrary sequence of
whitespace and a means to normalize it to a single space.  Using this framework,
we can generate non-bijective lenses using our framework, as long as the
canonized forms of their source and target types are in bijective
correspondence.

For Category 2 non-bijectivity, we could compose the generated lens (that has
projected information at the end) with one that deletes that information at the
tail of the file, when going left to right, and restores that information when
going back. This wrapper is not difficult to write, and we think that
discovering this projection automatically is possible, but we have not yet done
it.

@@

The preliminaries define a specific subset of Boomerang that we aim to
synthesize.  Boomerang supports a variety of lenses - asymmetric lenses,
quotient lenses, and alignment lenses defined over a series of papers
[Combinators for bi-directional tree transformations: a linguistic approach to
the view update problem, POPL 2005; Boomerang: Resourceful Lenses for String
Data, POPL 2008; Quotient Lenses, ICFP 2008].  In the preliminaries, we define
which subset of Boomerang we aim to synthesize. While this language does not
appear verbatim in a prior paper, it is a natural core language for string
lenses.

                                -------------

     - p.4, "naively using it in the context of lens synthesis presents several
       problems"? Which problems? At this point I have no clue what "using it"
       may even mean. The "big picture" is not clear at all.

The big picture is: we are trying to carve out a core, bijective subset of
Boomerang to synthesize.  We want to find a subset that is
sufficiently expressive, but also permits an efficient synthesis algorithm.
For this subset to be sufficiently expressive, it is important to have a type
equivalence rule, where a regular expression type can be substituted for an
equivalent one.  However, we choose to use a notion of equivalence that is finer
than language equivalence: star semiring equivalence.  The reasons we choose to
use star semiring equivalence instead of language equivalence are detailed on
pages 8 through 10.

### 2017-09-13 papers/BijectiveLensSynth/popl_18_rebuttal.txt

One more technical note about point: FlashFill's DSL does permit nested
loops, but for efficiency reasons the algorithm only synthesizes
programs with no nesting (this is discussed at the end of section 4.4 of
"Automating String Processing in Spreadsheets Using Input-Output
Examples", POPL '11).  While a single nesting of loops seems to be
sufficient for FlashFill's intended uses, our benchmark suite (based on
real-world uses of the Augeas tool) includes many more complex formats and
transformations.  In particular, these formats tend to have deep layers of
iterated data, with complex transformations at each layer, requiring the
synthesis of many layers of nested loops.  FlashFill cannot handle such
examples.

                                -------------

     - What modifications did we need to make to examples in the Augeas and
       FlashFill benchmark suite to make them bijective?

Many reviewers expressed concern that major changes were added to the formats in
our benchmark suite to make them bijective. We made changes to the benchmark
formats to handle two forms of non-bijectivity:
1) Whitespace was present in one format but not the other.
2) Useful information was projected away going from one format to the other.

@@

Category 2 was less prevalent, but still present, typically in the FlashFill
examples.  To address this category, we added additional information to the end
of the target format containing the projected information.

                                -------------

     - The restriction to bijective lenses is not fully justified

There are many formats kept in bijective correspondence with each other, like
those present at https://www.data.gov/ .

@@

See general responses.

                                -------------

     - Important parts of algorithm not explained (regarding how examples are
       used)

We omitted the formal description of these orderings for space.  A full
description of the algorithm is present in algorithm 4 in the appendix (see the
supplemental pdf submitted with the paper) on page 119, with the orderings
defined in Definition 33 on page 118 (the definitions of related structures
begin at page 115).  We embed the parse trees of the examples into the DNF
regular expressions, and make an ordering on the DNF regular expression with
embedded parse trees.  We would be happy to rewrite the paper to include
whatever information the reviewer believes is most important.

@@

                                -------------

     - I also have a concern: since the approach is
       limited to bijective lenses, but many of the test examples were not
       bijective, how much work was it to modify the examples
       and how easy is it to manually hack the resulting bijective
       lenses to make them work on the real data?  Is there any chance of
       automating this or generalizing to handle non-bijective lenses
       (perhaps with this algorithm as a subroutine)?

See general responses.

For Category 1 non-bijectivity, instead of changing the formats, we could have
placed canonizers on the edges of formats with large amounts of whitespace.
Canonizers can normalize an arbitrarily long sequence of whitespace to a single
space.  Canonizers also can be used to handle casing issues: normalizing a
sequence of uppercase and lowercase characters into purely uppercase characters.
In this manner, whitespace (or casing) information can be normalized before
synthesis, and no longer would need to be propagated to the target format. We
have recently developed a framework to annotate a regular expression with
canonizing information, allowing for a single term to represent an arbitrary
sequence of whitespace and a means to normalize it to a single space.  Using
this framework, we easily can generate non-bijective lenses using our framework,
as long as the canonized forms of their source and target types are in bijective
correspondence.

@@

Number of expansions is the correct phrase.  Figure 9 shows how much random
search through expansions is required before RigidSynth terminates.  In
particular, this shows that full expansion inference performs significantly
better than the partial one that only performs the expansions it knows it must
perform.

                                -------------

     - p21.  Also interesting to evaluate would be how time / number of
       solutions varies (e.g. for the full algorithm) as the number of
       examples provided varies.  Is the number of examples small?

The number of examples provided doesn't typically have an impact on runspeed,
only correctness.  The number of valid solutions does vary with the number of
examples provided.  Typically, a small number of examples can greatly reduce the
number of possible solutions.  The number of examples is indeed small, as is
shown in Figure 11.

For example, there are an average of 2709 solutions with 0 randomly generated
examples, 2016 with 2 randomly generated examples, and 755 with 5 randomly
generated examples.  We are omitting the data from aug/bbhosts, whose number of
possible solutions (in the 10^17 range) overwhelmed all other data.

@@

We can make sure we can provide at a description of the sizes of specs and
lenses in the final version of the paper.

                                -------------

     - Very limited DSL for transformations. The proposed DSL can only describe
       transformations that use identities and constant functions over input
       subparts. For example, one cannot synthesize the function toUpperCase,
       which I imagine could be useful for the kind of transformations
       appearing in the motivation.

We can convert between all uppercase and all lowercase.  Quotient lenses can be
used to normalize a format (or portions of a format) into fully uppercase or
fully lowercase data.  We believe quotient lenses have a bijective core, so our
synthesis engine for bijective lenses can be used as a plug-in component for
tasks (like to_upper) requiring quotient lenses.

@@

See general response.

                                -------------

     - On this note, there was a paper this year at PLDI on inverting string
       encoders (Automatic program inversion using symbolic transducers, by Hu
       et al.), a topic which seems very related, but for which I couldn’t find
       a citation.

Our DSL contains transformations that cannot be expressed by Extended Symbolic
Finite Transducers.  The finite lookahead does not allow for transformations
that swap data of indeterminite size, a common pattern and important pattern in
Boomerang lenses. We will add a citation to this work and a comment detailing
the restrictions of extended symbolic finite transducers.

@@

Boomerang doesn't currently provide such capabilities.  However, there is no
technical reason this can't be done by adding in a reverse-and-iterate
operator to Boomerang.  Only a slight adjustment to the synthesis algorithm
would be needed to incorporate this into the algorithm.

                                -------------

     - What is the difference in size between the regexes for domain/range spec
       and the full lenses?

On average, the domain and range specs are slightly larger than the lens itself.
See above for the chart and a commentary explaining why we believe writing
lenses is harder than writing regular expressions, despite the fact that lenses
alone are sometimes shorter than regular expression specifications.

### 2017-09-13 papers/BijectiveLensSynth/popl_18_rebuttal.txt

                                -------------

     - What is the difference in expressiveness between our work and
        existing algorithms like FlashFill?

Formally, they are incomparable: each can synthesize some transformations
that the other cannot.

Since we restrict ourselves to bijections, we are unable to express functions
that
don't preserve data, such as
  - mapping from a format with whitespace to one without whitespace,
  - turning a mix of uppercase and lowercase characters into purely
    uppercase characters, and
  - extracting a file name from a file path.
FlashFill can express all these.

@@

Conversely, our "Optician" algorithm can solve some classes of problems that
existing synthesis systems cannot.  In particular, techniques that only use
input-output examples are unable to synthesize programs with nested loops.

One more technical note about point: FlashFill's DSL does permit nested
loops, but for efficiency reasons the algorithm only synthesizes
programs with no nesting (this is discussed at the end of section 4.4 of
"Automating String Processing in Spreadsheets Using Input-Output
Examples", POPL '11).  While a single nesting of loops seems to be
sufficient for FlashFill's intended uses, our benchmark suite (based on
real-world uses of the Augeas tool) includes many more complex formats and
transformations.  In particular, these formats tend to have deep layers of
iterated data, with complex transformations at each layer, requiring the
synthesis of many layers of nested loops.  FlashFill cannot handle such
examples.  Our experiments demonstrate this fact by showing that FlashFill
only succeeds on [x] out of our [y] benchmarks, whereas our system succeeds on
all our benchmarks.

@@

     - The restriction to bijective lenses is not fully justified

There are many formats kept in bijective correspondence with each other, like
those present at https://www.data.gov/ .

There are lens theories developed to handle synchronization tasks for data
formats not kept in bijective correspondence (quotient lenses to handle Category
1 non-bijectivity, and symmetric lenses to handle Category 2 non-bijectivity).
These theories have a bijective core, surrounded by either canonizers (for
quotient lenses) or projections (for symmetric lenses).  We believe that our
system can be used as a drop-in component for synthesis algorithms on these more
complex lens structures.

Reviewer 1 asked us directly about the potential for future work.
Along those lines, we can say we have already implemented a system that
synthesizes quotient lenses from regular expressions augmented with equivalence
relations that, for instance, say that whitespace is irrelevant.  This system
uses the algorithm provided in this paper as a drop-in component without change.
We conjecture the system could be augmented in a similar way with projections as
well.  We are not suggesting adding more results of this form to the paper
because the paper is already very dense, but transformations of the form
(f o bijection o g) are very powerful, and can be synthesized using our engine
as a drop-in component when f and g are provided.

@@

                                -------------

     - After having written the regular expressions, is it very difficult
       for users to then write the lens?  Are we saving anybody any
       significant effort?

Engineering best practices dictate that we annotate the lenses with their
types.  While Boomerang does not need these annotations - being able to
infer the types from the terms - the types of the term serve as
documentation for future programmers to understand what formats the lens
maps between.  Furthermore, it provides an additional resilience to future
modifications of the lens, as it validates that the lens maps exactly
between strings of the provided formats - invalid inputs are not accepted
nor are the outputs ever invalid.  In this way, we are not requiring
additional work by the programmer to use our system; we are allowing the
programmer to save the difficulties of writing the term by using work they
should already be performing.

There are a few more reasons why programming with regular expressions are easier
than programming with lenses.

* Regular expressions are common - almost every programmer has a firm
grip on regular expressions.  In contrast, only a handful of programmers know
anything about programming with lenses.

* A regular expression describes one piece of data.  When writing a lens, one
must think about both pieces of data at the same time, and simultaneously
translate back and forth between them.  This is a non-standard programming task.

* One must sometimes compute in their own head how to unroll or transform the
regular expression into the right form, before writing the lens.

TODO: move to appropriate places in other users, and remove *s from 18B
To highlight these difficulties, we show an
example program for mapping the author field of a BibTeX entry:

@@

This is correct.

                                -------------

     - p19: "easy to understand" - can you present any evidence for this?
       Easy for whom?

We feel that they are easy to understand because they look similar to how we
would handwrite the lenses, but we have not done any user studies to justify
that they are easy to understand.  Evidence to the fact that they are easier to
understand than non-pretty printed versions is given by the sizes of the
generated lenses.  Without minimization, the lenses sizes average at 424 AST
nodes, whereas minimization made the sizes average at 276 AST nodes.
In the general response, we show what a minimized vs handwritten lens look like.
Below we show the difference between a non-minimized title field and a minimized
title field for a work item.

### 2017-09-13 papers/BijectiveLensSynth/popl_18_rebuttal.txt

Formally, they are incomparable: each can synthesize some transformations
that the other cannot.

Since we restrict ourselves to bijections, we are unable to express functions
that
don't preserve data, such as
  - mapping from a format with whitespace to one without whitespace,
  - turning a mix of uppercase and lowercase characters into purely
    uppercase characters, and
  - extracting a file name from a file path.
FlashFill can express all these.

Conversely, our "Optician" algorithm can solve some classes of problems that
existing synthesis systems cannot.  In particular, techniques that only use
input-output examples are unable to synthesize programs with nested loops.

One more technical note about point: FlashFill's DSL does permit nested
loops, but for efficiency reasons the algorithm only synthesizes
programs with no nesting (this is discussed at the end of section 4.4 of
"Automating String Processing in Spreadsheets Using Input-Output
Examples", POPL '11).  While a single nesting of loops seems to be
sufficient for FlashFill's intended uses, our benchmark suite (based on
real-world uses of the Augeas tool) includes many more complex formats and
transformations.  In particular, these formats tend to have deep layers of
iterated data, with complex transformations at each layer, requiring the
synthesis of many layers of nested loops.  FlashFill cannot handle such
examples.  Our experiments demonstrate this fact by showing that FlashFill
only succeeds on 3 out of our 39 benchmarks, whereas our system succeeds on
all our benchmarks.

@@

===========================================================================
Review #18B

                                -------------

     - However, I’m not sure what exact advantage the proposed approach brings
       to the table. In particular, the user is still required to write exact
       input and output types of the transformations via regular expressions.
       How hard is to write lenses after one has such expressions? ...
       In the reviewer’s opinion, the specification mechanism is not very
       natural and it’s unclear whether it actually simplifies writing
       bijective transformations. This aspect is not evaluated and it would be
       nice to at least see the difference in sizes between the domain/range
       regexes vs the Boomerang lenses. In general, there is no report of the
       sizes of benchmarks.

See general response.  We provide an example program (and specification) below:

      typedef WSP = (" " | "\t" | "\n")+;;
      typedef NAME = UPPERCASE (LOWERCASE)*;;

We also provide data on the sizes below:

@@

===========================================================================
Review #18C

                                -------------

     - Overall, this paper looks like a solid and useful contribution to the
       lens literature. However, this reviewer has never used, let alone
       designed, a lens library. As such, I found the paper very hard to read
       and assess. I would have liked to see a clear problem statement (such as
       an example where it's annoying to write the Boomerang code but easy to
       synthesize the lens using your approach

We provide an example program below.

      typedef WSP = (" " | "\t" | "\n")+;;
      typedef NAME = UPPERCASE (LOWERCASE)*;;

We will make the problem statement more clear in the final version.

@@

We will make the problem statement more clear in the final version.

                                -------------

     - why is there no Boomerang code for your example in Sec. 3?).

We can include example code like we did above to highlight the frustrations a
lens programmer must encounter while writing their programs.

### 2017-09-13 papers/BijectiveLensSynth/popl_18_rebuttal.txt

Reviewer 1 asked us directly about the potential for future work.
Along those lines, we can say we have already implemented a system that
synthesizes quotient lenses from regular expressions augmented with equivalence
relations that, for instance, say that whitespace is irrelevant.  This system
uses the algorithm provided in this paper as a drop-in component without change.
We conjecture the system could be augmented in a similar way with projections as
well.  We are not suggesting adding more results of this form to the paper
because the paper is already very dense, but transformations of the form
(f o bijection o g) are very powerful, and can be synthesized using our engine
as a drop-in component when f and g are provided.

     - After having written the regular expressions, is it very difficult
       for users to then write the lens?  Are we saving anybody any
       significant effort?

Engineering best practices dictate that we annotate the lenses with their
types.  While Boomerang does not need these annotations - being able to
infer the types from the terms - the types of the term serve as
documentation for future programmers to understand what formats the lens
maps between.  Furthermore, it provides an additional resilience to future
modifications of the lens, as it validates that the lens maps exactly
between strings of the provided formats - invalid inputs are not accepted
nor are the outputs ever invalid.  In this way, we are not requiring
additional work by the programmer to use our system; we are allowing the
programmer to save the difficulties of writing the term by using work they
should already be performing.

A few more reasons why programming with regular expressions is easier
than programming with lenses:

* Regular expressions are widely understood.  Few programmers know about
  programming with lenses.

* A regular expression describes one data format, while lenses deal with two
  at the same time.

* In the (common) case where the two formats have kleene stars in different
  places, manual lens programming requires mentally transforming both to a
  common "aligned" form.  Our synthesis algorithm does this alignment
  automatically.

### 2017-09-13 papers/BijectiveLensSynth/popl_18_rebuttal.txt

Formally, they are incomparable: each can synthesize some transformations
that the other cannot.

Since we restrict ourselves to bijections, we are unable to express functions
that
don't preserve data, such as
  - mapping from a format with whitespace to one without whitespace,
  - turning a mix of uppercase and lowercase characters into purely
    uppercase characters, and
  - extracting a file name from a file path.
FlashFill can express all these.

Conversely, our "Optician" algorithm can solve some classes of problems that
existing synthesis systems cannot.  In particular, techniques that only use
input-output examples are unable to synthesize programs with nested loops.

One more technical note on this point: FlashFill's DSL does permit nested
loops, but for efficiency reasons the algorithm only synthesizes
programs with no nesting (this is discussed at the end of section 4.4 of
"Automating String Processing in Spreadsheets Using Input-Output
Examples", POPL '11).  While a single nesting of loops seems to be
sufficient for FlashFill's intended uses, our benchmark suite (based on
real-world uses of the Augeas tool) includes many more complex formats and
transformations.  In particular, these formats tend to have deep layers of
iterated data, with complex transformations at each layer, requiring the
synthesis of many layers of nested loops.  FlashFill cannot handle such
examples.  Our experiments demonstrate this fact by showing that FlashFill
only succeeds on 3 out of our 39 benchmarks, whereas our system succeeds on
all our benchmarks.

### 2017-09-13 papers/BijectiveLensSynth/popl_18_rebuttal.txt

Formally, they are incomparable: each can synthesize some transformations
that the other cannot.

Since we restrict ourselves to bijections, we are unable to express functions
that
don't preserve data, such as
  - mapping from a format with whitespace to one without whitespace,
  - turning a mix of uppercase and lowercase characters into purely
    uppercase characters, and
  - extracting a file name from a file path.
FlashFill can express all these.

Conversely, our "Optician" algorithm can solve some classes of problems that
existing synthesis systems cannot.  In particular, techniques that only use
input-output examples are unable to synthesize programs with nested loops.

We note that FlashFill's DSL does permit nested
loops, but for efficiency reasons the algorithm only synthesizes
programs with no nesting (this is discussed at the end of section 4.4 of
"Automating String Processing in Spreadsheets Using Input-Output
Examples", POPL '11).  While a single nesting of loops seems to be
sufficient for FlashFill's intended uses, our benchmark suite (based on
real-world uses of the Augeas tool) includes many more complex formats and
transformations.  In particular, these formats tend to have deep layers of
iterated data, with complex transformations at each layer, requiring the
synthesis of many layers of nested loops.  FlashFill cannot handle such
examples.  Our experiments demonstrate this fact by showing that FlashFill
only succeeds on 3 out of our 39 benchmarks, whereas our system succeeds on
all our benchmarks.

### 2017-09-13 papers/BijectiveLensSynth/popl_18_rebuttal.txt

Conversely, our "Optician" algorithm can solve some classes of problems that
existing synthesis systems cannot.  In particular, techniques that only use
input-output examples are unable to synthesize programs with nested loops.

We note that FlashFill's DSL does permit nested
loops, but for efficiency reasons the algorithm only synthesizes
programs with no nesting (this is discussed at the end of section 4.4 of
"Automating String Processing in Spreadsheets Using Input-Output
Examples", POPL '11).  While a single nesting of loops seems to be
sufficient for FlashFill's intended uses, our benchmark suite (based on
real-world uses of the Augeas tool) includes many more complex formats and
transformations.  In particular, these formats tend to have deep layers of
iterated data, with complex transformations at each layer, requiring the
synthesis of many layers of nested loops.  FlashFill cannot handle such
examples.  Our experiments demonstrate this fact by showing that FlashFill
only succeeds on 3 out of our 39 benchmarks, whereas our system succeeds on
all our benchmarks.

     - What modifications did we need to make to examples in the Augeas and
       FlashFill benchmark suite to make them bijective?

Reviewers expressed concern that major changes were added to the formats in
our benchmark suite to make them bijective. We made changes to the benchmark
formats to handle two forms of non-bijectivity:
1) Whitespace was present in one format but not the other.
2) Useful information was projected away going from one format to the other.

Category 1 was by far the most common, applying to many of the Augeas
examples.  To handle it, we added additional whitespace to the other format,
where typically whitespace is unnecessary.

Category 2 was less prevalent, but present, typically in the FlashFill
examples.  To address this category, we added additional information to the end
of the target format containing the projected information.

### 2017-09-13 papers/BijectiveLensSynth/popl_18_rebuttal.txt

Conversely, our "Optician" algorithm can solve some classes of problems that
existing synthesis systems cannot.  In particular, techniques that only use
input-output examples are unable to synthesize programs with nested loops.

We note that FlashFill's DSL does permit nested
loops, but for efficiency reasons the algorithm only synthesizes
programs with no nesting (this is discussed at the end of section 4.4 of
"Automating String Processing in Spreadsheets Using Input-Output
Examples", POPL '11).  While a single nesting of loops seems to be
sufficient for FlashFill's intended uses, our benchmark suite (based on
real-world uses of the Augeas tool) includes many more complex formats and
transformations.  In particular, these formats tend to have deep layers of
iterated data, with complex transformations at each layer, requiring the
synthesis of many layers of nested loops.  FlashFill cannot handle such
examples.  Our experiments demonstrate this fact by showing that FlashFill
only succeeds on 3 out of our 39 benchmarks, whereas our system succeeds on
all our benchmarks.

     - What modifications did we need to make to examples in the Augeas and
       FlashFill benchmark suite to make them bijective?

Reviewers expressed concern that changes were made to the formats in
our benchmark suite to make them bijective. We made changes to handle two forms of non-bijectivity:
1) Whitespace was present in one format but not the other.
2) Useful information was projected away going from one format to the other.

### 2017-09-13 papers/BijectiveLensSynth/popl_18_rebuttal.txt

Conversely, our "Optician" algorithm can solve some classes of problems that
existing synthesis systems cannot.  In particular, techniques that only use
input-output examples are unable to synthesize programs with nested loops.
We note that FlashFill's DSL does permit nested
loops, but for efficiency reasons the algorithm only synthesizes
programs with no nesting (this is discussed at the end of section 4.4 of
"Automating String Processing in Spreadsheets Using Input-Output
Examples", POPL '11).

While a single nesting of loops seems to be sufficient for FlashFill's intended 
uses, our benchmark suite (based on
real-world uses of the Augeas tool) includes many more complex formats and
transformations.  In particular, these formats tend to have deep layers of
iterated data, with complex transformations at each layer, requiring the
synthesis of many layers of nested loops.  FlashFill cannot handle such
examples.  Our experiments demonstrate this fact by showing that FlashFill
only succeeds on 3 out of our 39 benchmarks, whereas our system succeeds on
all our benchmarks.

     - What modifications did we need to make to examples in the Augeas and
       FlashFill benchmark suite to make them bijective?

We changed some formats in our benchmark suite to handle two forms of non-bijectivity:
1) Whitespace present in one format but not the other.
2) Useful information projected away when going from one format to the other.

@@

Category 1 was by far the most common, applying to many of the Augeas
examples.  To handle it, we added additional whitespace to the other format,
where typically whitespace is unnecessary.

Category 2 was less prevalent, but occurred, typically in the FlashFill
examples.  To address this category, we added the projected information to the end
of the format missing the information.

### 2017-09-14 papers/BijectiveLensSynth/popl_18_rebuttal.txt

     - The restriction to bijective lenses is not fully justified

There are many formats kept in bijective correspondence with each other, like
those present at https://www.data.gov/ .

There are lens theories developed to handle synchronization tasks for data
formats not kept in bijective correspondence (quotient lenses to handle Category
1 non-bijectivity, and symmetric lenses to handle Category 2 non-bijectivity).
These theories have a bijective core, surrounded by either canonizers (for
quotient lenses) or projections (for symmetric lenses).  We believe that our
system can be used as a drop-in component for synthesis algorithms on these more
complex lens structures.

Reviewer 1 asked us directly about the potential for future work.
Along those lines, we have already implemented a system that
synthesizes quotient lenses from regular expressions augmented with equivalence
relations that solve the whitespace bijection problem.  This system
uses the algorithm provided in this paper as a drop-in component without change.
We conjecture the system could be augmented in a similar way with projections as
well.  We are not suggesting adding more results of this form to the current paper
because it is already very dense.

### 2017-09-14 papers/BijectiveLensSynth/popl_18_rebuttal.txt

There are lens theories developed to handle synchronization tasks for data
formats not kept in bijective correspondence (quotient lenses to handle Category
1 non-bijectivity, and symmetric lenses to handle Category 2 non-bijectivity).
These theories have a bijective core, surrounded by either canonizers (for
quotient lenses) or projections (for symmetric lenses).  We believe that our
system can be used as a drop-in component for synthesis algorithms on these more
complex lens structures.

Reviewer 1 asked us directly about the potential for future work.
Along those lines, we have already implemented a system that
synthesizes quotient lenses from regular expressions augmented with equivalence
relations that solve the whitespace bijection problem.  This system
uses the algorithm provided in this paper as a drop-in component without change.
We conjecture the system could be augmented in a similar way with projections as
well.  We are not suggesting adding more results of this form to the current paper
because it is already very dense.

     - After having written the regular expressions, is it very difficult
       for users to then write the lens?  Are we saving anybody any
       significant effort?

Engineering best practices dictate that we annotate the lenses with their
types.  While Boomerang does not need these annotations - being able to
infer the types from the terms - the types of the term serve as
documentation for future programmers to understand what formats the lens
maps between.  Furthermore, the types provide resilience in the face of future
lens modifications, as they validate that the lens maps exactly
between strings of the provided formats. Invalid inputs are not accepted
nor are the outputs ever invalid.  In this way, we are not requiring
additional work; we are saving programmers from the difficulties of 
writing the lens by leveraging the work they should already be performing.

### 2017-09-14 papers/BijectiveLensSynth/popl_18_rebuttal.txt

There are lens theories developed to handle synchronization tasks for data
formats not kept in bijective correspondence (quotient lenses to handle Category
1 non-bijectivity, and symmetric lenses to handle Category 2 non-bijectivity).
These theories have a bijective core, surrounded by either canonizers (for
quotient lenses) or projections (for symmetric lenses).  We believe that our
system can be used as a drop-in component for synthesis algorithms on these more
complex lens structures.

Reviewer 1 asked us directly about the potential for future work.
Along those lines, we have already implemented a system that
synthesizes quotient lenses from regular expressions augmented with equivalence
relations that solve the whitespace bijection problem.  This system
uses the algorithm provided in this paper as a drop-in component without change.
We conjecture the system could be augmented in a similar way with projections as
well.  We are not suggesting adding more results of this form to the current paper
because it is already very dense.

     - After having written the regular expressions, is it very difficult
       for users to then write the lens?  Are we saving anybody any
       significant effort?

Engineering best practices dictate that we annotate the lenses with their
types.  While Boomerang does not require these annotations - being able to
infer the types from the terms - the types of the term serve as
documentation for future programmers to understand what formats the lens
maps between.  Furthermore, the types provide resilience in the face of future
lens modifications, as the types ensure that the lens maps exactly
between strings of the provided formats. Invalid inputs are not accepted
nor are the outputs ever invalid.  Consequently, we are not requiring
additional work. Instead, we are saving programmers from having to 
writing the lens by leveraging work they should already be performing.
